---
title: "PDR Clothing and Item Response Theory"
subtitle: "Analysis"
author: "Robert Chapman"
date: "12/05/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Data Prep and Description**  
```{r Codebook_Loading_and_Data_Prep}

#osf_download(osf_ls_files(osf_retrieve_node("https://osf.io/eyn3x/"))[3,], path="data/", conflict="overwrite")

source("data/Generate_Codebook.R")

#Codebook_list is a list object with this structure: 
#"Description" - string of description of the variable
#"Value" - vector of numeric values for the data, e.g., 1-5
#"Labelled" - vector of labelled response options, e.g., "never", "sometimes", "always"

#osf_download(osf_ls_files(osf_retrieve_node("https://osf.io/eyn3x/"))[1,], path="data/", conflict="overwrite")
Data=read.csv("data/engeln_zola_2021.csv")

```
1. Data Processing and Cleaning
    + Visual inspection of data for clear errors or empty rows/columns
      - *Base R: View()*
```{r View_Data, eval=FALSE, echo=TRUE} 
View(Data) #12/5: reviewed - no obvious issues with data 
```
    + Checking that variables are correctly formatted 
      - *Base R: class()*
```{r Check_Classes, eval=TRUE, echo=TRUE}
  ###Are all items and response formatted correctly - e.g., numeric v factor v value labels
  sapply(Data, class) #need to manually review for factor and labels, etc.
#12/5: all integer or numeric, which should be correct
```
    + Cross-check data variables with data dictionary
      - *Base R: names(Data) %in%  names(Codebook)*
```{r Names_Check, eval=TRUE, echo=TRUE}
  ###Are all items present?
  if(any(!(names(Data) %in% names(Codebook_list)))){
    print(paste("names in Data, but not in Codebook:",names(Data)[which(!(names(Data) %in% names(Codebook_list)))]))
    print(paste("names in Codebook, but not in Data:",names(Codebook_list)[which(!(names(Codebook_list) %in% names(Data)))]))
  }else{print("All names in Data and Codebook match")}
#12/5: "All names in Data and Codebook match"
```
    + Checking that variables meet expected range/values
      - *Base R: unique(Data[,var]) %in% Codebook[Var]$Values*
```{r ExpectedRangesValues, eval=TRUE, echo=TRUE}
#Out of Range values 
  sapply(names(Codebook_list), function(Var){
    if(!all(is.na(Codebook_list[[Var]]$Value))){
      unique(Data[,Var])[!(unique(Data[,Var]) %in% Codebook_list[[Var]]$Value)]}})
#12/5: All variables look fine, but freq items (e.g., shoepainfreq), contain 6 levels, not 5. Will alert authors, but no changes needed here- since I won't be using the frequency items.
```
    + Visualization
      - *Base R: Boxplots(), pairs()*  
```{r PlotPairs, eval=TRUE, echo=TRUE}
#reducing for the variables used in analysis
PDRitems=names(Data)[!grepl("obcs|freq|bas|gender|sample|bodysurv",names(Data))]

# Barplots
source("data/Outlier_Detection.R")
Barplots(Data[,PDRitems])
#12/5: Done with real data

#Pairs Plots
jpeg("pairs.jpg", width=2200, height=2200)
par(mfrow=c(length(PDRitems),length(PDRitems)))
for(i in PDRitems){for(j in PDRitems){
barplot(table(Data[,i], Data[,j]), main=paste(i,j))}}
dev.off()
#12/5: Done with real data

#Reviewed- no obvious problems, easier to read a correlation matrix than it is the pairs plots of staked bar charts.

```
2. Missing Data
    + Evaluate missingness patterns/Littles MNAR test/T test comparisons
      - *‘Missing Data’ script*  
```{r MissingData, eval=TRUE, echo=TRUE}
source("data/Missing_Data.R")

Missing_Patterns(Data[,PDRitems[!grepl("skirts|breasts",PDRitems)]])
Missing_Patterns(Data[which(Data$gender==1),PDRitems])
Missing_Patterns(Data[which(Data$gender==2),PDRitems])
#12/5: not enough missingness (~1-4 cases for each missingness pattern distributed across items) to do gender by missingness comparison- it is worth noting that 4 males are missing the 'bend' item and no females are missing that item. Maybe worth taking a look at descriptive stats by gender for that item


```
3. Normality  
    + Skew/Kurt/QQplot
      - *‘Assumption Checking’ script*
```{r AssumptionChecking, eval=TRUE, echo=TRUE}
source("Assumption_Checking.R")
SK=sapply(Data[,PDRitems], SkewKurtosis)
which(abs(SK["Skew",])>2)
which(abs(SK["Kurtosis",])>2)
#12/5:only the 'breasts' items violates the kurtosis recommendations (>2), and by a hair (-2.0029). Unlikely to cause an issue, and will be retained.

QQplots(Data[,PDRitems])
#12/5: QQplots were generated - arms, breaths, drag and walk not have a great QQ lines - but that may be due to just lower frequency of endorsement, sub QQ with binomial 


```

    + Outlier detection
      - Boxplot/zRange/plausible meanSD’s
        + *‘Outlier Detection’ script*  
```{r OutlierDetection, eval=TRUE, echo=TRUE}
source("data/Outlier_Detection.R")

boxplot(apply(Data[,PDRitems],1,sum,na.rm=T)) #12/5: 1 value of sum score 13 is outside the IQR

zRange(apply(Data[,PDRitems],1,sum,na.rm=T)) #12/5: z range = -3.1 to 2.26, but the sum score of 13 is inside of the z score range of |3.29|, 3.1- retain this case for now. May delete later if it appears that this case isn't a part of the population, but it doesn't appear that way now. 

apply(Data[,PDRitems],2, meanSD)#12/5: all item-level means/SDs appear reasonable mean 1.32:1.85 & sd 0.36:0.50

```

3. Descriptives  
    + Mean, Median, Standard Deviation, Frequency, Range
      - *Base R: mean(), median(), sd(), table(), range())*
```{r MeanMedianSDFreqRange, eval=TRUE, echo=TRUE}

rbind(
"All"=apply(Data[,PDRitems], 2, meanSD),
"Male"=apply(Data[which(Data$gender==1),PDRitems], 2, meanSD),
"Female"=apply(Data[which(Data$gender==2),PDRitems], 2, meanSD))
#12/5:Done

rbind(
"All"=apply(Data[,PDRitems], 2, median, na.rm=T),
"Male"=apply(Data[which(Data$gender==1),PDRitems], 2, median, na.rm=T),
"Female"=apply(Data[which(Data$gender==2),PDRitems], 2, median, na.rm=T))
#12/5:Done

write.csv(t(rbind(
"All"=apply(Data[,PDRitems], 2, function(x){paste0(meanSD(x),", ",median(x, na.rm = TRUE))}),
"Male"=apply(Data[which(Data$gender==1),PDRitems], 2, function(x){paste0(meanSD(x),", ",median(x, na.rm = TRUE))}),
"Female"=apply(Data[which(Data$gender==2),PDRitems], 2, function(x){paste0(meanSD(x),", ",median(x, na.rm = TRUE))}))),"PDRitemByGender_MeanSDMedian.csv")


sapply(Data[,PDRitems], table)#12/5:Done

sapply(Data[,PDRitems], table, Data$gender)#12/5:Done

write.csv(t(rbind(sapply(Data[,PDRitems], table),sapply(Data[,PDRitems], table, Data$gender))),"PDRItems_Frequency_TotalByGender.csv" )

range(apply(Data[,PDRitems], 1, sum, na.rm=T))#12/5:Done, sum-score range= 13-30
```
    + Correlation matrix 
      - *psych package: tetrachoric()*
```{r CorMat, eval=TRUE, echo=TRUE}
#install.packages("psych")
library(psych)
corr=round(tetrachoric(Data[,PDRitems])$rho,2)#12/5:Done
range(corr[lower.tri(corr)]) #12/5:rho's range from 0.11:0.73, seems reasonable
hist(corr[lower.tri(corr)]) #12/5: most corr coefs are around 0.3-0.5, again, reasonable
write.csv(corr,"PDRitems_TetrachoricCorr.csv")

```

    + Standardized Cronbach’s Alpha
      - *psych package: alpha()*    
```{r Alpha, eval=TRUE, echo=TRUE}
alpha(Data[,PDRitems])$total$std.alpha #12/5:Done - 0.83 alpha internal consistency
```

## Replication of paper analyses
  
1. Dimensionality
    + Scree plots
      - *psych package: fa.parallel()*
```{r ScreePlots, eval=FALSE, echo=TRUE}
Scree_total=fa.parallel(Data[,PDRitems[!grepl("skirts|breasts",PDRitems)]])$fa.values
Scree_male=fa.parallel(Data[which(Data$gender==1),PDRitems[!grepl("skirts|breasts",PDRitems)]])$fa.values
Scree_female=fa.parallel(Data[which(Data$gender==2),PDRitems])$fa.values
plot(Scree_total, type="l", ylab="Eigenvalues", xlab="Factor Number")
lines(Scree_male, col="red")
lines(Scree_female, col="blue")
legend(8, 3.25, legend=c("All", "Male", "Female"),
       col=c("black","red", "blue"), lty=1, cex=0.7)
```
    + ICLUST
      - *psych package: ICLUST()*
```{r ICLUST, eval=FALSE, echo=TRUE}
ICLUST(tetrachoric(Data[,PDRitems[!grepl("skirts|breasts",PDRitems)]])$rho)
ICLUST(tetrachoric(Data[which(Data$gender==1),PDRitems[!grepl("skirts|breasts",PDRitems)]])$rho)
ICLUST(tetrachoric(Data[which(Data$gender==2),PDRitems])$rho)
```

    + Bi-factor modeling
      - *psych package: omega()*  
```{r Omega, eval=FALSE, echo=TRUE}
omega(tetrachoric(Data[,PDRitems[!grepl("skirts|breasts",PDRitems)]])$rho)
omega(tetrachoric(Data[which(Data$gender==1),PDRitems[!grepl("skirts|breasts|breaths",PDRitems)]])$rho)
omega(tetrachoric(Data[which(Data$gender==2),PDRitems])$rho)
###breaths had to be deleted from male bifactor model due to an ultra-heywood case. This was justified due to a small sample size of men endorsing this item (smallest #, less than 10%)
```

*review of dimensionality analyses results alongside items and literature review to qualitative guide decision on number of factors to included or modelled*
  
2. IRT Assumption Checks  
	*Will need to be further fleshed out for which model is selected*
    + Monotonicity
          - *assumed for dichotomous data*
      - Unidimensionality
        + Omega hierarchical/total & Explained Common Variance
          - *psych package: omega()*
```{r OmegaHOmegaTotECV, eval=FALSE, echo=TRUE}
omega(tetrachoric(Data[,PDRitems])$rho)[c("omega_h", "omega.tot", "ECV")]
```
      - Local dependence 
        + *Rasch: Q3*
        + *2pl: jackknife slope index – Q3  for rasch*
      - Item invariance (DIF)
        + *Rasch: Mantel-Haenszel*
        + *2PL: logistic ordinal regression for 2pl?*
3. IRT Modeling
    + Models to be evaluated: 
      - Rasch 
        - *psych package: rasch()* 
      - Rasch w/guessing 
        - *???psych package: rasch()* 
      - 2PL 
        - *psych package: ltm()* 
      - Multidimensional IRT 
        - *mirt package: mirt()* 
      - IRTree Modeling 
        - *???? package: ???()* 
      - Explanatory IRT Modeling 
        - *???? package: ???()* 
    + Model fit
      - Goodness of fit/deviance 
        - *Package: AIC,BIC, Chi-Sq?*
    + Item fit
      - Infit/outfit/signed chisq
        - *Package: TBD*  
  
## What are the criteria for confirming and disconfirming the hypotheses?
  
### Hypothesis 1  
There is either a unidimensional or multidimensional construct of ‘discomfort/distraction for appearance’ as measured by the PDR clothing items?  
  
**Confirming**  
1. There is additional literature supporting a construct of ‘discomfort/distraction for appearance’ as indicated by wearing PDR clothing.  
2. Scree plots, ICLUST and Bifactor analyses indicate a reasonable number of factors (1 to 5) in the PDR clothing set
  
**Disconfirming**  
1. There is additional literature contradicting a construct of ‘discomfort/distraction for appearance’ as indicated by wearing PDR clothing.  
2. Scree plots, ICLUST and Bifactor analyses indicate an unreasonable or untestable number of factors (e.g., 6+) in the PDR clothing set
  
### Hypothesis 2  
An Item Response Theory model can be fit to the dimensions in the PDR clothing item data, and will allow further evaluation and comparison of PDR clothing items in relation to the central construct(s) of ‘discomfort/distraction for appearance’.  
  
**Confirming**  
1. Model assumptions will be met for each IRT model to be fit, including monotonicity, local dependence and item invariance  
2. Goodness of fit indices for the model is appropriate (e.g., CFI/TLI >0.95, RMSEA <0.08) and for item fit (e.g., non-significant signed chi-square)  
  
**Disconfirming**  
1. Model assumptions are not met for each IRT model to be fit, including monotonicity, local dependence and item invariance  
2. Goodness of fit indices for the model are not appropriate (e.g., CFI/TLI <0.95, RMSEA >0.08) and item fit (e.g., significant signed chi-square)  

## Have the analyses been validated on a subset of the data?

The analyses have not been validated on a subset of the data 

## What is known about the data that could be relevant for the tested hypotheses?

Original publication introducing the PDR item displayed high internal consistency (cronbach’s alpha >0.9) and group-level differences between gender.

## Please provide a brief timeline for the different steps in the preregistration?

Planning and writing for the preregistration started in October, 2021. Data was simulated based on the codebook published on OSF [https://osf.io/ajv5z/](https://osf.io/ajv5z/) in November, 2021 and published on GitHub under a branch of the "Prereg_Function" project, called "ClothingIRT" - (https://github.com/zenit125/Prereg_Functions/tree/Clothing_IRT). Set-up and hypothesis building will completed in December with results of additional literature review and simulated data/coding embedded in this parent RMarkdown document. The actual research below is planned for completion in December 2021.
  
Hypothesis 1 will be tested after completion of the set-up and hypothesis building. The results of Hypothesis 1 will be posted in an expanded version of this RMarkdown document.  
  
Based on the results of Hypothesis 1 and any requisite additional background research, revisions to Hypothesis 2, coding and results will occur and will be documented in this parent RMarkdown document. 
